---
title: 'SDGB 7844 HW 1: R Syntax and Control Structures'
author: 'Name: Jiayin Hu &nbsp; &nbsp; Class Time: Thu. 1:15-3:15PM'
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Question 1

The vectors `state.name`, `state.area`, and `state.region` are pre-loaded in R and contain US state names, area (in square miles), and region respectively. 

First, have a look at the content of these 3 vectors.

```{r}
state.name
state.area
state.region
```


(a) Identify the data type for `state.name`, `state.area`, and `state.region`.

```{r}
typeof(state.name)
typeof(state.area)
typeof(state.region)
```
The data type of these three vectors are character, double and integer respectively.

(b) What is the longest state name (including spaces)? How long is it?

```{r}
#Using if statement
state.name.len = 0 # store the longest length
y <- c() #store the order number of the state with longest name

for (i in 1:length(state.name)){
  x <-  nchar(state.name[i])   #the length of each state
  if (state.name.len < x){     
    state.name.len = x
    y <- i
  }
  else{
    if (state.name.len == x) y <- c(y, i)
  }
}

print(state.name[y])  #output the state name
state.name.len
```
The state "North Carolina" and "South Carolina" have the longest length. They both contains 14 bytes.

(c) Compute the average area of the states which contain the word “New” at the start of the state name. Use the function `substr()`.

```{r}
state.area.total <- 0 #initial sum of area
new.state <- c()      #store the state beginning with "New"

for (i in 1:length(state.name)){
  if(substr(state.name[i], start = 1, stop = 3) == "New"){
    new.state <- c(new.state, state.name[i])
    state.area.total <- state.area.total + state.area[i]
  }
}

print(new.state)
state.area.average <- state.area.total/length(new.state)
print(paste("The average area of the states which contain the word “New” at the start of the state name is", state.area.average))
```

(d) Use the function `table()` to determine how many states are in each region. Use the function `kable()` to include the table in your solutions. (Notes: you will need the R package _knitr_ to be able to use `kable()`. See the RMarkdown example in the Assignments folder on Blackboard for an example.)

```{r}
library(knitr)
state.table <- table(state.region)
kable(state.table, caption = "State Region Table", align = "c" )
```


##Quesetion 2

Perfect numbers are those where the sum of the proper divisors (i.e., divisors other than the number itself) add up to the number. For example, 6 is a perfect number because its divisors, 1, 2, and 3, when summed, equal 6.

(a) The following code was written to find the first 2 perfect numbers: 6 and 28; however, there are some errors in the code and the programmer forgot to add comments for readability. Debug and add comments to the following:

```{r}
num.perfect <- 2  #to find 2 perfect numbers
count <- 0  #the number of the perfect number we've found
iter <- 2  #the number used to check whether it is a perfect number or not. The initial value is 2. 

while(count < num.perfect){  # modify <= to <. When there's less than 2 numbers, keep going
  
  divisor <- 1  #1 is a divisorof all numbers.
  
#Use for circulation to find all the divisors except the number itself.
  for(i in 2:(iter-1)){
    
    if(iter%%i==0) divisor <- c(divisor, i)
    
  } # end for loop

#Determine whether the sum of the dicisors equals to the number.
#If true, then output the number is a perfect number and add 1 to count.
  if(sum(divisor)==iter){     # modify = to ==
    
    print(paste(iter, " is a perfect number", sep=""))
    count <- count + 1 
    
  } # end if
  
  iter <- iter + 1 #check the next number in the next circulation
  
} # end while loop
```


(b) Use the function `date()` at the start and at the end of your amended code. Then compute how long the program approximately takes to run (you can do this sub- traction by hand). Find the run time when you set `num.perfect` to 1, 2, 3, and 4. Create a table of your results (Note: use the first table format in the RMarkdown Example file in the Assignments folder on Blackboard) . What are the first four perfect numbers?

```{r}

for (t in 1:4){
  print(date())
  num.perfect <- t  #to find t perfect numbers
  count <- 0  #the number of the perfect number we've found
  iter <- 2  #the number used to check whether it is a perfect number or not. The initial value is 2. 
  
  while(count < num.perfect){  #When there's less than t numbers, keep going
  
    divisor <- 1  #1 is a divisorof all numbers.
  
  #Use for circulation to find all the divisors except the number itself.
    for(i in 2:(iter-1)){
    
      if(iter%%i==0) divisor <- c(divisor, i)
    
    } # end for loop
  
  #Determine whether the sum of the dicisors equals to the number.
  #If true, then output the number is a perfect number and add 1 to count.
    if(sum(divisor)==iter){     
    
      print(paste(iter, " is a perfect number", sep=""))
      count <- count + 1 
    
    } # end if
  
    iter <- iter + 1 #check the next number in the next circulation
  
  } # end while loop
  print(date())
}
```
So the first 4 perfect number is 6, 28, 496, 8128. 
The rough run time is shown as follow. 
```{r, echo = FALSE}
timeData <- data.frame("No of Perfect Numbers"=c(1, 2, 3, 4), "Run Time"=c(0, 0, 0, 8))
kable(timeData, caption = "Run Time Table", align = "c" )
```

(c) Let x <- 1:4 and define y to be the vector of run times. Plot y vs x using the code below. Is the relationship between the discovery of perfect numbers and run times on your computer linear? Justify your answer.

```{r}
x <- 1:4
y <- c(0, 0, 0, 8)
plot(x, y, pch=20, type="b", xlab="number of perfect numbers discovered",
     ylab="cumulative length of time (in seconds)", 
     main="Cumulative Run Times to Discover Perfect Numbers", las=TRUE, cex.main = 1.0)
```

The relationship between the discovery of perfect numbers and run times on your computer is not linear. Run time increases significantly when the number we aim to find is larger than 3, because we have to run more than 8000 loops.

Here the time is roughly calculated by hands. In another way, we use `Sys.time()` to record the time and repeat (b) and (c). Meanwhile, this function can subtract directly to get the run time. 

```{r}
y <- c()
for (t in 1:4){
  start <- Sys.time()
  num.perfect <- t  #to find n perfect number
  count <- 0  #the number of the perfect number we've found
  iter <- 2  #the number used to check whether it is a perfect number or not. The initial value is 2. 
  
  while(count < num.perfect){  #When there's less than n numbers, keep going
  
    divisor <- 1  #1 is a divisorof all numbers.
  
  #Use for circulation to find all the divisors except the number itself.
    for(i in 2:(iter-1)){
    
      if(iter%%i==0) divisor <- c(divisor, i)
    
    } # end for loop
  
  #Determine whether the sum of the dicisors equals to the number.
  #If true, then output the number is a perfect number and add 1 to count.
    if(sum(divisor)==iter){     # ==
    
      print(paste(iter, " is a perfect number", sep=""))
      count <- count + 1 
    
    } # end if
  
    iter <- iter + 1 #check the next number in the next circulation
  
  } # end while loop
  end <- Sys.time()
  time <- end - start
  y <- c(y, time)
  print(time)
}
x <- 1:4
timeData <- data.frame("No of Perfect Numbers"= x, "Run Time"=y)
kable(timeData, caption = "Run Time Table", align = "c" )
plot(x, y, pch=20, type="b", xlab="number of perfect numbers discovered",
     ylab="cumulative length of time (in seconds)", 
     main="Cumulative Run Times to Discover Perfect Numbers", las=TRUE, cex.main = 1.0)
```

##Question 3

(a) Using a for loop, write code to compute the geometric mean of the numeric vector x < −c(4, 67, 3). Make sure your code (i) removes any NA values and (ii) prints an error message if there are any non-positive values in x.

```{r}
geoMean <- function(x){
  
  #remove any NA values
  xWithoutNa <- x[!is.na(x)]
  n <- length(xWithoutNa)
  product <- 1
  
  # print an error message if there are any non-positive values in x.
  if (any(xWithoutNa<=0)) cat("Error: (", x, ") contains non-positive value \n")
  
  else {
    for (i in 1:n){
      product <- product * xWithoutNa[i]
      geometricMean <- product^(1/n)
    }
  cat("The geometric mean of (", x, ") is", geometricMean, "\n")
  }
}

x <- c(4, 67, 3)
geoMean(x)
```


(b) Test your code on the following cases and show the output: (i) {NA,4,67,3}, (ii) {0, NA, 6}, (iii) {67, 3, ∞}, and (iv) {−∞, 67, 3}.
```{r}

x1 <- c(NA,4,67,3)
x2 <- c(0, NA, 6)
x3 <- c(67, 3, Inf)
x4 <- c(-Inf, 67, 3)
for (i in 1:4){
  geoMean(eval(parse(text = paste0("x", i))))
  
}

```


