---
title: "SDGB 7840 Homework 4"
author: "Name: Jiayin Hu Class Time: 3:30-5:30pm"
date: "2019/4/25"
output: pdf_document
---
Before the homework, we list the library we need first.
```{r message=FALSE, warning=FALSE}
require(tidyverse)
require(knitr)
require(ggplot2)
require(gridExtra)
require(GGally)
require(readxl)
require(pROC)
require(usdm)
```
### Logistic Regression

Read data.
```{r, message=FALSE}
date <- read_csv("SpeedDating.csv", col_names=TRUE)
```

#Q1
```{r}
# Contingency table
xtabs(~DecisionM + DecisionF, data = date)
nrow(date)
```
From the contingency table, we can see that the two people in 63 dates want a second date. The percentage is 
$$\frac{63}{276}=22.83\%.$$

#Q2
```{r}
# Make a new column 'second.date'
date['second.date'] <- 0
date[(date$DecisionM == 1) & (date$DecisionF == 1), 'second.date'] <- 1
```

```{r}
theme.info <- theme(plot.title = element_text(size=10, hjust=0.5),
                    axis.title = element_text(size=10),
                    axis.text = element_text(size=10),
                    legend.position = 'top')
# Jitter plot
j_like <- date %>% ggplot(aes(x = LikeM, y=LikeF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Rate of Liking") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

j_partneryes <- date %>% ggplot(aes(x = PartnerYesM, y=PartnerYesF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Probability that partner says \"yes\"") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

s_age <- date %>% ggplot(aes(x = AgeM, y=AgeF, col=factor(second.date))) +
  geom_point(size=2) + ggtitle("Age") +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

j_attractive <- date %>% ggplot(aes(x = AttractiveM, y=AttractiveF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Rate of Attractiveness") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

j_sincere <- date %>% ggplot(aes(x = SincereM, y=SincereF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Rate of Sincerity") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

j_intelligent <- date %>% ggplot(aes(x = IntelligentM, y=IntelligentF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Rate of Intelligence") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

j_fun <- date %>% ggplot(aes(x = FunM, y = FunF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Rate of Fun") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

j_ambitious <- date %>% ggplot(aes(x = AmbitiousM, y=AmbitiousF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Rate of Ambition") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

j_interest <- date %>% ggplot(aes(x = SharedInterestsM, y=SharedInterestsF, col=factor(second.date))) +
  geom_point(size=2, position = "jitter") + ggtitle("Rate of Shared Interests") +
  scale_x_continuous(breaks=seq(0, 10, 1)) + 
  scale_y_continuous(breaks=seq(0, 10, 1)) +
  scale_color_manual(values=c("#ef528580", "#35386680")) +
  labs(x="Male", y="Female") + theme.info

```
```{r warning=FALSE, fig.height = 10, fig.width = 10}
jitter <- grid.arrange(s_age, j_like, j_partneryes, j_attractive, j_sincere, j_intelligent, 
                        j_fun, j_ambitious, j_interest, nrow=3)
```
The outcomes (second date or no second date) are not completely separable for each numerical variable. It is obvious that the observations of second date tend to rate higher for liking and attractiveness.


#Q3
```{r}
# check reasonableness of numerical variables
for (col in colnames(date[, c(3:6, 11:22)])){
  r <- any(date[col]<1 | date[col] > 10, na.rm = TRUE)
  if (r){
    print(paste("Not all responses to ", col, " are within range from 1 to 10."))
    row <- which(date[col]<1 | date[col] > 10)
    # print(row)
    print(date[row, col])
  }else{
    print(paste("All responses to ", col, " are within range from 1 to 10."))
  }
}
```
According to the output, the responses to PartnerYesM, FunM, SharedInterestsM, SharedInterestsF are not within the range 1-10. After inspecting each variable, all the values without the range are 0. These should be treated as wrong records need correcting. These maybe caused by the respondents' cursoriness ignoring the range is from 1. So we correct the 0 answer to 1.

```{r}
datec = date %>% mutate(PartnerYesM = replace(PartnerYesM, PartnerYesM == 0, 1)) %>%
  mutate(FunM = replace(FunM, FunM == 0, 1)) %>%
  mutate(SharedInterestsM = replace(SharedInterestsM, SharedInterestsM == 0, 1)) %>%
  mutate(SharedInterestsF = replace(SharedInterestsF, SharedInterestsF == 0, 1))
```

```{r}

# check missing data in numerical variables

# datec[, c(3:6, 11:22)] %>% summarize_all(funs(sum(is.na(.))))
# able to use, does not show well in rmarkdown

for (col in colnames(datec[, c(3:6, 11:22)])){
  #exist_na <- any(is.na(datec[col])==TRUE)
  count_na <- sum(is.na(datec[col])==TRUE)
  if (count_na>0){
    print(paste(as.character(count_na), " missing observation(s) in responses to ", col))
  }else{
    print(paste("No missing observations in responses to ", col))
  }
}
```
Missing data exist in the responses and the output above shows the specific number of missing data. 

#Q4
```{r}
# check race categories
unique(c(datec$RaceM, datec$RaceF))
unique(datec$RaceM)
unique(datec$RaceF)
```
The possible race categories are Caucasian, Asian, Latino, Black and Other, for both male and famale.
```{r}
# check missing data in race
racem_count_na <- sum(is.na(datec["RaceM"]) == TRUE) # 2
racef_count_na <- sum(is.na(datec["RaceF"]) == TRUE) # 4
racemf_count_na <- sum((is.na(datec["RaceM"])) | (is.na(datec["RaceF"])) == TRUE) # 6
racem_count_na
racef_count_na
racemf_count_na
```
2 missing data in race of male and 4 missing data in race of famale. There are 6 records in total missing race information in the datset. It is hard to amend the missing race information. If `RaceF`, `RaceM` are included in the model, the missing value should be removed. 

```{r fig.height=6, fig.width=10}
mosaicplot(table(datec$RaceM, datec$RaceF), main = "Race Categories", 
           xlab = "Male Race", ylab = "Female Race", 
           las = TRUE, cex.axis =1, color=c("#67D5B595", "#EE778595", "#C89EC495", "#84B1ED95", "#DE7E7395"))

# mosaicplot(table(datec$RaceF, datec$RaceM), main = "Race Categories", 
#            xlab = "Female Race", ylab = "Male Race", 
#            las = TRUE, cex.axis =1, color=c("#67D5B595", "#EE778595", "#C89EC495", "#84B1ED95", "#DE7E7395"))
```
Most participants are Caucasian and Asian males and females. The most dates are between Caucasian male and Caucasian female. No Black females date with Black males and other race males. 

#Q5

```{r, fig.height=10, fig.width=10, results="hide"}
# Model selection
complete_date <- datec[complete.cases(datec),-c(1, 2)]

model.full <- glm(second.date ~ ., data = complete_date, family = binomial(link="logit"))

step <- step(model.full, direction="backward", trace = 1, test="Chisq")

model.attemp1 <- glm(second.date ~ LikeM + FunF + PartnerYesM + PartnerYesF + AgeM + 
                    RaceF + AttractiveF + AmbitiousF, data=datec, family = binomial(link="logit"))
summary(model.attemp1)

model.attemp13 <- glm(second.date ~ LikeM + FunF + PartnerYesM + PartnerYesF +  
                       AttractiveF + AmbitiousF, data=datec, family = binomial(link="logit"))
summary(model.attemp13)
```
To select the model, we first only keep complete cases in the dataset and use backward stepwise method to find possible explanatory variables in the final model. However, when we establish the model with those possible explanatory variables, we only exclude those observations that have missing values in the variables that are actually included in that final model.[^1] 

After backward stepwise method based on AIC, we obtain `model.attemp1` with `LikeM `, `FunF`, `PartnerYesM`, `PartnerYesF`, `AgeM`, `RaceF`, `AttractiveF`, `AmbitiousF`. After that we remove the insiginificant variables and obtain `model.attemp13` as `model.final` with explanatory variables `LikeM `, `FunF`, `PartnerYesM`, `PartnerYesF`, `AttractiveF`, `AmbitiousF`.

[^1]: The process refers https://rcompanion.org/rcompanion/e_07.html 


```{r}
# final model
model.final <- glm(second.date ~ LikeM + FunF + PartnerYesM + PartnerYesF +  
                    AttractiveF + AmbitiousF, data = datec, family = binomial(link="logit"))
summary(model.final)
```

In this dataset, the responses of explanatory variables are from respondents and we have corrected the irrational values, so that they should be measured without error. 

The model seems corretly specified since the 6 variables includes rating questions for both male and female and the coefficients are reasonable according to our intuition. 

From the scatterplots in Q2, we can see that outcomes are not completely separable for each explanatory variables.

The plot of Cook's distance does not show obvious outliers in the dataset.

Each record is collected from each pair of participants, so the independence of observations should be ok.

Multicollinearity problem can be excluded, since VIF values for all explanatory variables in our final model are less than 10.

```{r}
date.final <- datec %>% dplyr::select(second.date, LikeM, FunF, PartnerYesM, PartnerYesF, AttractiveF, AmbitiousF) %>% na.omit()
plot(cooks.distance(model.final), , type="h", las=TRUE, main="Cook's Distance")
usdm::vif(as.data.frame(date.final[-c(1)]))
```

```{r}
# Can also be done by calculating from summary(model.final)
model.null <- glm(second.date ~ 1, data=date.final, family = binomial(link="logit"))
#anova(model.final, model.null, test="Chisq")
anova(model.final, model.null, test="LRT")
```
Likelihood ratio test for overall model:

$$H_0: \beta_{LikeM} = \beta_{FunF} = \beta_{PartnerYesM} = \beta_{PartnerYesF} = \beta_{AttractiveF} = \beta_{AmbitiousF} = 0$$
$$H_a: \text{At least one of the slopes is not 0.}$$
The test statistic $G=95$ which has a $\chi^2$ distribution with $df=6$. The $p$-value is $2.2 \times 10^{-16}$, which is less than $\alpha=0.05$. Therefore, we reject the null hypothesis and the overall model is significant.

z-test for slopes:
$$H_0: \beta_{LikeM} = 0 \quad \text{v.s.} H_a: \beta_{LikeM} \neq 0$$
The test statistic $z=3.673$ which has a standard normal distribution. The $p$-value is $0.000239$, which is less than $\alpha=0.05$. Therefore, we reject the null hypothesis and $\beta_{LikeM}$ is statistically significant.

$$H_0: \beta_{FunF} = 0 \quad \text{v.s.} H_a: \beta_{FunF} \neq 0$$
The test statistic $z=2.406$ which has a standard normal distribution. The $p$-value is $0.0161$, which is less than $\alpha=0.05$. Therefore, we reject the null hypothesis and $\beta_{FunF}$ is statistically significant.

$$H_0: \beta_{PartnerYesM} = 0 \quad \text{v.s.} H_a: \beta_{PartnerYesM} \neq 0$$
The test statistic $z=3.321$ which has a standard normal distribution. The $p$-value is $0.0008$, which is less than $\alpha=0.05$. Therefore, we reject the null hypothesis and $\beta_{PartnerYesM   }$ is statistically significant.

$$H_0: \beta_{PartnerYesF} = 0 \quad \text{v.s.} H_a: \beta_{PartnerYesF} \neq 0$$
The test statistic $z=2.592$ which has a standard normal distribution. The $p$-value is $0.0095$, which is less than $\alpha=0.05$. Therefore, we reject the null hypothesis and $\beta_{LikeM}$ is statistically significant.

$$H_0: \beta_{AttractiveF} = 0 \quad \text{v.s.} H_a: \beta_{AttractiveF} \neq 0$$
The test statistic $z=2.361$ which has a standard normal distribution. The $p$-value is $0.0182$, which is less than $\alpha=0.05$. Therefore, we reject the null hypothesis and $\beta_{AttractiveF}$ is statistically significant.

$$H_0: \beta_{AmbitiousF} = 0 \quad \text{v.s.} H_a: \beta_{AmbitiousF} \neq 0$$
The test statistic $z=-2.374$ which has a standard normal distribution. The $p$-value is $0.0176$, which is less than $\alpha=0.05$. Therefore, we reject the null hypothesis and $\beta_{AmbitiousF}$ is statistically significant.



#Q6
```{r}
date.new <- datec[complete.cases(datec %>% dplyr::select(second.date, LikeM, FunF, PartnerYesM, PartnerYesF, AttractiveF, AmbitiousF)),]
nrow(date.new)
xtabs(~DecisionM + DecisionF, data = date.new)
kable(table(date.new$second.date), col.names = c("second.date", "Freq"))
```

We have 6 explanatory variables in the final model. According to the rule of thumb, at least 10 observations for each outcome per predictor. Therefore we need at least 60 samples for each outcomes. In our model, we have 193 no second date and 62 second date. So the number of explanatory variables follows rule of thumb.

#Q7
Holding other variables constant, if rate of LikeM increases by 1, the odds of second date increases by $(e^{0.4940}-1)\times 100\% =$ `r (exp(0.4940)-1) *100`%.

Holding other variables constant, if rate of FunF increases by 1, the odds of second date increases by $(e^{0.3486}-1)\times 100\% =$ `r (exp(0.3486)-1) *100`%.

Holding other variables constant, if rate of PartnerYesM increases by 1, the odds of second date increases by $(e^{0.3416}-1)\times 100\% =$ `r (exp(0.3416)-1) *100`%.

Holding other variables constant, if rate of PartnerYesF increases by 1, the odds of second date increases by $(e^{0.2693}-1)\times 100\% =$ `r (exp(0.2693)-1) *100`%.

Holding other variables constant, if rate of AttractiveF increases by 1, the odds of second date increases by $(e^{0.2860}-1)\times 100\% =$ `r (exp(0.2860)-1) *100`%.

Holding other variables constant, if rate of AmbitiousF increases by 1, the odds of second date decreases by $(e^{0.3047}-1)\times 100\% =$ `r (exp(0.3047)-1) *100`%.

`LikeM`, `FunF`, `PartnerYesM`, `PartnerYesF`, `AttractiveF` increase the probability of a second date. This corresponds to our intuition that if you appreciate, understand your partner and the feeling between both you is good, a second date is more probable.

`AmbitiousF` decreases the probability of a second date. It is hard to tell ambition is good or not in an relationship. But an ambitious man maybe a little aggressive and proud in a speed dating, which may causes females feeling not so well. 


#Q8
```{r}
AUC <- auc(response=date.final$second.date, predictor=model.final$fitted.values)
ROC <- roc(response=date.final$second.date, predictor=model.final$fitted.values,
    plot=TRUE, las=TRUE, 	legacy.axes=TRUE, lwd=5,
    main="ROC for Second Date", cex.main=1, cex.axis=1, cex.lab=1)
legend("bottomright",legend=paste("AUC=", round(AUC, digits=3), sep=" "),
       bty="n", cex=1)
```

```{r}
temp <- as.data.frame(t(coords(ROC, x="best", ret=c("threshold", "specificity", "sensitivity"))))
kable(coords(ROC, x="best", ret=c("threshold", "specificity", "sensitivity")))

temp_df <- data.frame(date.final, "fitted.values"=round(model.final$fitted.values, digits=3))
classify.best <- rep(1, times=nrow(temp_df))
classify.best[temp_df$fitted.values < coords(ROC, x="best", ret="threshold")] <- 0

table(classify.best, temp_df$second.date)
```
The AUC is 0.872. And the best threshold for classifying observations is 0.1614, which means when estimated probability is equal or great than 0.1614, we will classify the observation as 1 and the observation with estimated probability less than 0.1614 will be classified as 0. 

```{r}
(131 + 56)/(131+56+62+6) # Accuracy
56/(56+6) # Sensitivity
131/(131+62) # Specificity
```

$$\text{Accuracy:} \quad \frac{131 + 56}{131+56+62+6} = 0.7333$$
$$\text{Sensitivity:} \quad \frac{56}{56+6} = 0.9032$$
$$\text{Specificity:} \quad \frac{131}{131+62} = 0.6788$$



### One-Way ANOVA
Read data
```{r}
kudzu_data <- read_excel("kudzu.xls", col_names = TRUE)
kudzu_data$Treatment <- factor(kudzu_data$Treatment, 
                               levels = c('Control', 'LowDose', 'HighDose'), 
                               ordered = TRUE)
```
#Q9

Response variable is bone mineral density in the femur ($\text{g/cm}^2$)

#Q10

Factors is the dosage of isoflavones. Levels are control group (no usage of isoflavones), a low dose of isoflavones and a high dose of isoflavones.

#Q11 

3 treatments are included, since only one factor is explored.

#Q12 

Since rats were randomly assigned and each group has 15 observations, the experimental design is a comletely randomized, balanced design.

```{r}
table(kudzu_data$Treatment) 
```

#Q13
```{r}
count = kudzu_data %>% group_by(Treatment) %>% count()
mean_sd = kudzu_data %>% group_by(Treatment) %>% summarize(mean(BMD), sd(BMD))
table = data.frame(full_join(count, mean_sd, by = "Treatment"))
kable(table, col.names = c( 'Treatment', "Sample Size", "BMD Mean ($\\text{g/cm}^2$)", "BMD Std ($\\text{g/cm}^2$)"), caption = 'Summary Table')
```

#Q14
```{r fig.height=5, fig.width=8}
# 14. side-by-side box plots
theme.info <- theme(plot.title = element_text(size=12, hjust=0.5),
                    axis.title = element_text(size=10),
                    axis.text = element_text(size=10))
kudzu_data %>% 
  ggplot(aes(Treatment, BMD, fill = Treatment)) + 
  geom_boxplot() +
  stat_summary(fun.y=mean, geom="line", aes(group=1), lwd=1, col="cadetblue") +
  stat_summary(fun.y=mean, geom="point", pch=19, size=1, col="firebrick") +
  ggtitle("Boxplots of BMD by Treatment Group\nWith Connected Means") +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  labs(x="Treatment Group", 
       y=expression(paste("BMD (", g/cm^{2},")"))) +
  theme.info + theme(legend.position="none")
```
From the box plot, the 1st quartile, 3rd quartile, median, maximum and mean BMD in HighDose group are obviously larger than those in Control group and LowDose group.

The 1st quartile, 3rd quartile, median and mean BMD in Control group and LowDose group are not much different. The mean and median BMD in LowDose group are a little lower than Control group.

#Q15

The experiment is a completely randomized design, so it satisfies independence.
And we have check the sample size in each treatment. This is a balanced design.

By the normal quantile plot of residuals, the points basically coincide with the theoratical line. So the residuals follow normal distribution.

The largest standard deviation is 0.0188 (HighDose) and the smallest standard deviation is 0.0115 (LowDose).According to the rule of thumb, since the largest standard deviation is less than twice the smallest standard deviation, it can be regarded as constant variance.

```{r fig.height=6, fig.width=6}
# residuals
mean <- kudzu_data %>%
  group_by(Treatment) %>%
  summarize(mean(BMD))
left_join(kudzu_data, mean, by = "Treatment") %>%
  mutate(residuals = BMD - `mean(BMD)`) %>%
  ggplot(aes(sample = residuals)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle("Normal Quantile Plot of Residuals") +
  theme.info
```

#Q16

One-way ANOVA hypotheses are:
$$H_0: \mu_{\text{Control}}=\mu_{\text{LowDose}}=\mu_{\text{HighDose}} \quad \text{v.s.} \quad  H_a: \text{At least two means are different}$$
The test statistic $F=7.718$ with 2 and 42 degrees of freedom. $p\text{-value}=0.0014<0.01=\alpha$. Therefore we reject null hypothesis, which means at least two means are different.
```{r}
# one-way ANOVA model
kudzu_anova <- aov(BMD ~ Treatment, data = kudzu_data)
summary(kudzu_anova)
```

#Q17
p-values of the difference in HighDose-Control, HighDose-LowDose are less than critical value 0.05. So HighDose group is significantly different from others. Also in the plot we can see that 0 is not in the confidence interval of those two pairwise differences. 
```{r, fig.width=10, fig.height=4}
# Tukey’s multiple-comparisons method
kudzu_Tukey <- TukeyHSD(kudzu_anova, conf.level = 0.95)
kudzu_Tukey
par(mar=c(2, 9, 3, 2))
plot(kudzu_Tukey, las=TRUE)
```
